---
title: "Additional Summaries of the Simulation in Section 4.1 of 'Efficient Estimation of the GP-P Regression Model'"
author: "Easton Huch and Candace Berrett"
date: "3/18/2022"
output: html_document
---

```{r setup, include=FALSE}

#NOTE: Code provided as-is

#Change to file path and data name
myfilepath <- "myfilepath"
myfilename <- "simulation41"

knitr::opts_chunk$set(echo = FALSE, warning=F)
library(tidyverse)
library(RColorBrewer)
library(knitr)
mypal <- brewer.pal(8, "Set2")
mypaldark <- brewer.pal(8, "Dark2")
load(paste(myfilepath, "/", myfilename, ".Rdata", sep=""))

#Create variables for values not saved in the file
Ps <- c(0.26, 1, 2, 3.4)
nreps <- 1000
nsamps <- length(sampsizes)
nphis <- length(phis)
nPs <- length(Ps)
modnames <- c("GP1", "GP2", "GP-0.26", "GP-3.4", "GP-mom", "nlm-T", "optim-T", "vglm-1", "vglm-2", "Pois", "QPois", "NB-1", "NB-2")
nmods <- length(modnames)
comptime[which(comptime==0)] <- NA 

#Create a data frame of all the saved values
mydat <- data.frame(comp1=c(comptime[1,,,,,]), comp2=c(comptime[2,,,,,]), comp2=c(comptime[3,,,,,]), betahat=c(betahat), betal=c(betalow), betau=c(betahigh), niters=c(niters), phihat=c(phihat), phil=c(philow), phiu=c(phihigh), P=factor(rep(rep(Ps, each=(nreps*nmods)), times=nphis*nsamps)), phi=factor(rep(phis, each=(nreps*nmods*nPs*nsamps))), nsize=factor(rep(rep(sampsizes, each=nreps*nmods*nPs), nphis)), mod=factor(rep(rep(modnames, each=nreps), times=nPs*nsamps*nphis), levels=modnames))
```


## $\hat{\beta}_1$

Boxplots of the 1000 estimates of $\hat{\beta}_1$ (y-axis) for the thirteen fitted models (x-axis) and three sample sizes (denoted by color) when data were generated from the GP-P model for various values of $P$ and $\varphi$.  The red horizontal line indicates the true data-generating value of $\beta_1$.  Model-fitting approaches that fit the ``true" data-generating model have a darker background.  The boxplots of estimates from the proposed maximum likelihood approach are slightly darkened.

```{r plot2betahat, fig.width=13, fig.height=10}
#Boxplots for beta-hat
myspacing <- c(0.75, .5, .25)
par(mfrow=c(4, 3), mar=c(1, 4, 4, 2))
for(thisP in Ps){
  for(thisphi in phis){
  plot(0:nmods, ylim=c(-1.5,3.5), xlim=c(0, nmods), xlab="Model", ylab=expression(hat(beta)), type='n', main=bquote(P == .(thisP) ~ phi == .(thisphi)), xaxt='n')
    rect(par("usr")[1], par("usr")[3],
     par("usr")[2], par("usr")[4],
     col = "#f7f7f7")
        abline(v=c(1,2,4,5,6,7,8,9), col='black')
    
    rect(4, -3, 7, 4, col=rgb(.9, .91, .93))   
    if(thisP==2){
      rect(1, -3, 2, 4, col=rgb(.9, .91, .93))
      rect(8, -3, 9, 4, col=rgb(.9, .91, .93))
    }
    if(thisP==1){
      rect(0, -3, 1, 4, col=rgb(.9, .91, .93))
      rect(7, -3, 8, 4, col=rgb(.9, .91, .93))
    }
    if(thisP==0.26){
      rect(2, -3, 3, 4, col=rgb(0.9, .91, .93))
    }
    if(thisP==3.4){
      rect(3, -3, 4, 4, col=rgb(.9, .91, .93))
    }
    abline(v=0:(nmods+1), lwd=1.5, col='white')
    abline(h=-2:3, col='white')
    abline(h=1, col='red', lwd=2)
    mycounter <- 1
    for(thisn in sampsizes){
         thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi),]
         boxplot(betahat~mod, add=T, at=(1:nmods)-myspacing[mycounter], data=thisdat, xaxt="n", border=mypal[mycounter], boxwex=0.2, cex=.25, col='white')
         if(thisP==0.26){
           thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP-0.26"),]
         }
         if(thisP==1){
           thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP1"),]
         }
         if(thisP==2){
           thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP2"),]
         }
         if(thisP==3.4){
           thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP-3.4"),]
         }
           boxplot(betahat~mod, add=T, at=(1:nmods)-myspacing[mycounter], data=thisdat, xaxt="n", border=mypaldark[mycounter], boxwex=0.2, cex=.25, col="white")
        
        mycounter <- mycounter + 1
    }
    axis(1, at=(1:nmods)-0.5, tick=F, labels=modnames, cex.axis=0.5, line=-1)
    if(thisP==0.26 & thisphi==-0.1){
      legend("topright", legend=paste("n =", sampsizes), col=mypal[1:3], lty=1, lwd=1, cex=.8)
    }
  }
}
```


## 95% CI on $\beta_1$: WIDTH AND COVERAGE

Median width of the 95\% confidence intervals on $\beta_1$ for the 1000 simulated data sets from the GP-P model with various values of $P$, $\varphi$, and $n$ for each of the 13 models fit.  
```{r betwidth, message=F}
temp <- mydat %>% 
   group_by(P, phi, nsize, mod) %>%
   summarise(width=round(median(betau - betal, na.rm=T), 3)) %>%
   pivot_wider(names_from=mod, values_from=width)
kable(temp)
```

Coverage of the 95% confidence intervals on $\beta_1$.
```{r coverage, message=F}
temp <- mydat %>% 
   group_by(P, phi, nsize, mod) %>%
   summarise(cover=round(mean(betal < 1 & betau > 1, na.rm=T), 3)) %>%
   pivot_wider(names_from=mod, values_from=cover)
kable(temp)
```

## COMPARISON OF ESTIMATION METHOD OF $\varphi$

Estimates of $\varphi$ when fit via maximum likelihood (y-axis) vs method of moments (x-axis) with the appropriate GP-P model (i.e., P set to true data-generating value).  The color indicates the sample size for the data used to estimate the model.  The vertical and horizontal lines mark the true data-generated value of $\varphi$.  The dashed diagonal line indicates when the estimates would be the same.  

```{r phiest, fig.width=13, fig.height=11}
par(mfrow=c(4, 3), mar=c(4, 4, 4, 2))
for(thisP in Ps){
  for(thisphi in phis){
      plot(mydat$phihat, mydat$phihat, xlab="", ylab="", type='n', main=bquote(P == .(thisP) ~ phi == .(thisphi)), xlim=c(thisphi - abs(thisphi)*1.5, thisphi+abs(thisphi)*1.5), ylim=c(thisphi - abs(thisphi)*1.5, thisphi+abs(thisphi)*1.5))
    if(thisphi==-0.1){
      mtext("Maximum Likelihood", side=2, line=2.5, cex=.9)
    }
    if(thisP==3.4){
      mtext("Moment", side=1, line=2.5, cex=.9)
    }
    rect(par("usr")[1], par("usr")[3],
     par("usr")[2], par("usr")[4],
     col = "#f7f7f7")
    lines(c(thisphi - abs(thisphi)*1.5, thisphi+abs(thisphi)*1.5), c(thisphi - abs(thisphi)*1.5, thisphi+abs(thisphi)*1.5), col='lightgray', lty=2, lwd=2)
    abline(h=thisphi, col='white', lwd=2)
    abline(v=thisphi, col='white', lwd=2)

    mycounter <- 1
    for(thisn in sampsizes){
          if(thisP==0.26){
            thisdat.mle <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP-0.26"),]
          }
         if(thisP==1){
           thisdat.mle <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP1"),]
         }
         if(thisP==2){
           thisdat.mle <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP2"),]
         }
         if(thisP==3.4){
           thisdat.mle <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP-3.4"),]
         }
        thisdat.mom <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP-mom"),]
        points(thisdat.mom$phihat, thisdat.mle$phihat,  cex=.25, col=mypal[mycounter], pch=19)
        mycounter <- mycounter + 1
    }
    if(thisP==0.26 & thisphi==-0.1){
      legend("topleft", legend=paste("n =", sampsizes), col=mypal[1:3], pch=19, cex=.8)
    }
  }
}
```



## COMPUTATION TIME 

Boxplots of the 1000 log computation times (y-axis) for the thirteen fitted models (x-axis) and three sample sizes (denoted by color) when data were generated from the GP-P model for various values of $P$ and $\varphi$.  Model-fitting approaches that fit the ``true" data-generating model have a darker background.  The boxplots of the computation time from the proposed maximum likelihood approach are slightly darkened. 

```{r plot2comptime, fig.width=13, fig.height=10}
#Boxplots for computation time
myspacing <- c(0.75, .5, .25)
par(mfrow=c(4, 3), mar=c(1,4,4,2))
for(thisP in Ps){
  for(thisphi in phis){
  plot(0:nmods, ylim=range(log(mydat$comp1), na.rm=T), xlim=c(0, nmods), xlab="Model", ylab=expression("log(computation time)"), type='n', main=bquote(P == .(thisP) ~ phi == .(thisphi)),  xaxt='n')
    rect(par("usr")[1], par("usr")[3],
     par("usr")[2], par("usr")[4],
     col = "#f7f7f7")
    rect(4, -8, 7, 4, col=rgb(.9, .91, .93))   
    if(thisP==2){
      rect(1, -8, 2, 4, col=rgb(.9, .91, .93))
      rect(8, -8, 9, 4, col=rgb(.9, .91, .93))
    }
    if(thisP==1){
      rect(0, -8, 1, 4, col=rgb(.9, .91, .93))
      rect(7, -8, 8, 4, col=rgb(.9, .91, .93))
    }
    if(thisP==0.26){
      rect(2, -8, 3, 4, col=rgb(0.9, .91, .93))
    }
    if(thisP==3.4){
      rect(3, -8, 4, 4, col=rgb(.9, .91, .93))
    }
    abline(v=0:(nmods+1), lwd=1.5, col='white')
    abline(h=seq(min(log(mydat$comp1), na.rm=T), max(log(mydat$comp1), na.rm=T), length=10), col='white')
    mycounter <- 1
    for(thisn in sampsizes){
         thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi),]
         boxplot(log(comp1)~mod, add=T, at=(1:nmods)-myspacing[mycounter], data=thisdat, xaxt="n", border=mypal[mycounter], boxwex=0.2, cex=.25, col='white')
         if(thisP==0.26){
           thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP-0.26"),]
         }
         if(thisP==1){
           thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP1"),]
         }
         if(thisP==2){
           thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP2"),]
         }
         if(thisP==3.4){
           thisdat <- mydat[which(mydat$P==thisP & mydat$nsize==thisn & mydat$phi==thisphi & mydat$mod=="GP-3.4"),]
         }
           boxplot(log(comp1)~mod, add=T, at=(1:nmods)-myspacing[mycounter], data=thisdat, xaxt="n", border=mypaldark[mycounter], boxwex=0.2, cex=.25, col="white")
        
         mycounter <- mycounter + 1
    }
    axis(1, at=(1:nmods)-0.5, tick=F, labels=modnames, cex.axis=0.5, line=-1)
        if(thisP==0.26 & thisphi==-0.1){
      legend("topright", legend=paste("n =", sampsizes), col=mypal[1:3], lty=1, lwd=1, cex=.8)
    }
  }
}
```


## NON-CONVERGENCE

The table below shows the proportion of times an algorithm did not converge for different data sets generated from the GP-P model for various values of $P$, $\varphi$, and sample sizes, $n$.  Algorithms for GP-P when $\varphi$ estimated via method of moments, maximization of the log-likelihood via "nlm" and "optim," and the NB-1 model sometimes did not converge.  All other algorithms/models were always able to fit the data.  

```{r checknas, message=F}
temp <- mydat %>% 
  group_by(P, phi, nsize, mod) %>%
   summarise(pna=mean(is.na(betahat))) %>%
   pivot_wider(names_from=mod, values_from=pna)
kable(temp)
```
